Identifying Values to Express Emotions with a Non-Anthropomorphic Platform

### 2016-12-06 JournalIJSR/Parts/Introduction.tex
Changes in international studies has been changed as it was informed. The IJSR: intro and related work was modify. It was added a HRI late breaking report that was submitted to HRI

%TODO Here, even before what follows, the need for emotion expression should be justified, ad possibily supported by literature

Emotions and mental states are not just presented through facial expression, but also by body postures and other features~\cite{Gelder2008}. However,  many psychological studies have been mostly focused on understanding the role of human face in emotion projection~\cite{Ekman2004},~\cite{kleinsmith2012affective} and mental states. This trend has been followed by the robotics community, where anthropomorphic faces (e.g.,~\cite{Arras2012},~\cite{Breazeal2002}) and bodies (e.g.,~\cite{Canamero2010},~\cite{haering2011},~\cite{Destephe2013}) have been widely used to convey emotions.
Nevertheless, in many situations the presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces~\cite{Breazeal2002} or limbs~\cite{Li2011}. In some cases, like, for instance, in floor cleaning robots, anthropomorphic characteristics could even be detrimental to robot's task accomplishment.
This generates the necessity to study other mechanisms that could help to project emotions, which could give people an idea about the robots' state, and engage the user in long term relations.

The amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010,Lakatos2014,Sharma2013,Novika2015}) remains still small in comparison to those that exploit anthropomorphic features. Moreover, these works do not prescribe any specific range of values for the characteristics to be used to express the  implemented emotionsb. For example, Suk and collaborators~\cite{NAM2014}, in their study, give specific values for acceleration, curvature, but their connection to specific emotions is not given. Rather, they give a relationship between their features and values in terms of valence and arousal. This could lead to select features to convey certain emotions, but it would be better to know the specific range of values for those features to convey the single emotions, or to know the values that could lead to misinterpret the emotional expression.

@@

This paper is organized as follows. The next section introduces previous studies done in human emotion and robot emotions projection. Section~\ref{sec:system} introduces the robotic platform and the software used in the experiment. The experiment's design is explained in section~\ref{sec:experiment}. Finally sections~\ref{sec:study} and~\ref{sec:result} present the study done and the obtained results.

>>>>

Social environments are intricate spaces where people behave in ways that lead to their acceptance in social groups. Similarly, studies in social robotics have shown that robots acceptance increases when robots project a high social presence~\cite{Heerink08}. The straightest way to fulfill is through imitation of humans' characteristics, such us body form, behaviors and social characteristics. This has inspired most researchers to focus on how express emotions and mental states exploiting anthropomorphic features, in many occasions relying just on faces. However emotions and mental states are not just presented through facial expression, but also by body postures and other features~\cite{Gelder2008}. Nonetheless in psychology has been a clear tendency to study the role of human face in emotion projection~\cite{Ekman2004},~\cite{kleinsmith2012affective} and mental states. This trend has been followed by robotics community, where anthropomorphic faces (e.g.,~\cite{Arras2012},~\cite{Breazeal2002}) and bodies (e.g.,~\cite{Canamero2010},~\cite{haering2011},~\cite{Destephe2013}) have been widely used to convey emotions.
Nevertheless, in many situations the presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces or limbs. In some cases, like, for instance, in floor cleaning robots, anthropomorphic characteristics could even be detrimental to robot's task accomplishment.
This generates the necessity to study other mechanisms that could help to project emotions, which could give people an idea about the robots' state, and engage the user in long term relations.

The amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010,Lakatos2014,Sharma2013,Novika2015}) remains still small in comparison to those that exploit anthropomorphic features. Moreover, these works do not prescribe any specific range of values for the characteristics to be used to express the  implemented emotions. For example, Suk and collaborators~\cite{NAM2014}, in their study, give specific values for acceleration, curvature, but their connection to specific emotions is not given. Rather, they gave a relationship between their features and values in terms of valence and arousal. Another possibility is the use of professional human actors to study how they convey certain emotions. However, a direct mapping between humans and robots is not possible~\cite{Saerbeck2007,Canamero2010} due to robots' physical capabilities. In addition, the significance of the agreement obtained from the studies that used human actors is on discussion~\cite{Russell2003}. As a consequence determine precise values to project emotions in non-anthropomorphic platforms is required.

@@

This paper is organized as follows. The next section introduces previous studies done in human emotion and robot emotions projection. Section~\ref{sec:system} introduces the robotic platform and the software used in the experiment. The experiment's design is explained in section~\ref{sec:experiment}. Finally sections~\ref{sec:experiment} and~\ref{sec:result} present the study done and the obtained results.

### 2016-12-06 JournalIJSR/Parts/Platform.tex
Changes in international studies has been changed as it was informed. The IJSR: intro and related work was modify. It was added a HRI late breaking report that was submitted to HRI

A holonomic platform was built to be used in the experiment. This type of platforms are characterized by the possibility to move in any direction without the necessity to have a specific orientation, i.e., they are free to move taking any desired orientation. Therefore, it was possible to imitate movements that are done by humans. For example, people do not have constaraints in movement, and can take any direction in any moment. The platform has a diameter of 25 cm and height of 25 cm. Figure~\ref{fig:ThirdDesign} shows the platform's blue prints and the real platform. Robot's frame of reference, in which all the velocities are going to be calculated, is depicted by the two black arrows. So, as it could be observed, to make the robot move forward a velocity along the $y$ axis is selected by the control system.

@@

%%%%%%%%%%%%%%%%%%%%%%%platform_base
\subsection{Software}
To guarantee that the correct robot's velocity for each motor could be achieved, a PID controller was implemented. PID's set point is established by a higher level controller that could receive two different types of commands. The first commands are robot's velocities ($<V_x,V_y,\omega>$), given in robot's reference frame. The second command is a set point ($<x, Y, \theta>$) to be reached, given in the world reference frame. This frame of reference is set every time the system is reset or boot. For example, if the robot finishes a trajectory and then it is reset, the new frame is going to be in the robot's current position.  %TODO From this description it seems that this second reference frame is again associated to the robot, in particular that it is a reference frame that is activated when the robot starts, and does not move with the robot. So it is a world reference frame reset each time the robot starts a new movement. It should have a name, to be found. If it was always associated with the robot it would have been called "deictic".

%TODO I'm sorry, but the following is an implementation detail that does not add anything to the scientific relevance of this paper, and this is not a paper about the interface.
%In addition to this low level control, a graphical interface, depicted in Figure~\ref{fig:experimental_interface}, was created to reduce the possibility of introducing wrong values for a desired sequence. This interface loads the sequences from a .txt file and displays sequences' numbers on it. Every time that a new sequence should be presented to a participant, the sequence's number is selected in the interface, which will display sequence's values. Once the robot has been positioned to the correct position, the execution could be started by clicking on send button. Here two commands are send, one resetting the controller and the second the desire position. In case that the sequence's execution should be aborted, it should be clicked the button stop.

%\begin{figure}
%	\centering
%	\includegraphics[width=0.45\textwidth]{./Images/ExperimentInterface.png} 
%	\caption{Interface used in the experiment. Once a sequence is selected, the interface shows sequence's values. Also, the interface give information about the current position of the robot and its velocity.}
%	\label{fig:experimental_interface}
%\end{figure}

>>>>

A holonomic platform was built to be used in the experiment. This type of platforms are characterized by the possibility to move in any direction without the necessity to have a specific orientation, i.e., they are free to move taking any desired orientation. Therefore, it was possible to imitate movements that are done by humans. For example, people do not have constaraints in movement, and can take any direction in any moment. The platform has a diameter of 25 cm and height of 25 cm. Figure~\ref{fig:ThirdDesign} shows the platform's blue prints and the real platform. Robot's frame of reference, in which all the velocities are going to be calculated, is depicted by the two black arrows. Therefore, as it could be observed, to make the robot move forward a velocity along the $y$ axis is selected by the control system.

@@

%%%%%%%%%%%%%%%%%%%%%%%platform_base
\subsection{Software}
To guarantee that the correct robot's velocity for each motor could be achieved, a PID controller was implemented. PID's set point is established by a higher level controller that could receive two different types of commands. The first commands are robot's velocities ($<V_x,V_y,\omega>$), given in robot's reference frame. The second command is a set point ($<x, Y, \theta>$) to be reached, given in the world reference frame. This world frame of reference is set every time the system is reset or boot. For example, if the robot finishes a trajectory and then it is reset, the new frame is going to be in the robot's current position.

### 2016-12-06 JournalIJSR/Parts/Results.tex
Changes in international studies has been changed as it was informed. The IJSR: intro and related work was modify. It was added a HRI late breaking report that was submitted to HRI

A table with subjects' answers was generated for each experience. Table~\ref{table:example_table} is an example of the generated tables.

@@

For each table, mean, standard deviation, and median were calculated. It was not possible to use ANOVA test over the data because the assumption of normality is not satisfied in the collected data. This was checked using the Shapiro-Wilk Test.
Additionally, a contingency table for each emotion was generated in each experience as it is depicted in Table~\ref{table:table_contingency}, where the intensity for the rest of emotions is calculated as their mean, including the option ``other''. For all tables, including the contingency,  the Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$) was calculated: it is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (e.g., $Kappa$), $\alpha$ is a generalization of several known reliability indices, and it applies to any number of observers, any number of categories, any type of data, incomplete or missing data, large and small sample sizes.

This calculation was done using the R's package \textit{irr}. To improve the table interpretation, it was decided to just register the emotions' alpha values that had a mean greater than zero. The $\alpha$ for each experience is shown in Figure~\ref{fig:alpha_experience}. Additionally, for each emotion was generated a table with a top ten ranking experiences that considers the following items:(i) the mean intensity of the respective emotions, (ii) the alpha agreement for the respective emotion, and (iii) the alpha agreement for the specific experience. Experiences in these tables were first organized by mean, then by the emotion's agreement and finally by general agreement. %TODO This is critical, it is not clear (make an example) and should be fully justified, since on this is based all the reasoning about results

The decision to give more importance to emotion's alpha was taken based on the situation that the participants could be agreed about the precise emotion but not for the rest of the words enlisted. %TODO You just said that the first is the mean, not alpha... This sentence is not clear, anyway
From the resulting tables~\ref{table:happy_top_ten},~\ref{table:Angry_top_ten},~\ref{table:Sad_top_ten}, and~\ref{table:Scared_top_ten}, it is possible to notice:

@@

	\item Fear was the only emotion that had six over ten movements obtaining both general and specific alpha agreement over 0.41, which is the lower bound for moderate agreement~\cite{Viera2005}. Also, it was mostly selected (eight over ten) when $direction = \pi$ and $orientation = \pi$. The angle and the angular velocity attributed six over ten times were $0.087$ and $2$, respectively. So it seems that people perceive the movement as a fear expression when the robot is looking at them and moving far from them fast.

	\item Sadness was mostly attributed to linear velocities of $0 mm/s$ (two over ten) and $200 mm/s$ (eight over ten), and angular velocities of $0 rad/s$ and $1 rad/s$, with higher predilection of the second (eight over ten). It seems that people attribute sadness to slow velocities with slow angular velocity and small oscillation angle. Regarding the other two features, there is not any defined pattern that could lead to make a generalization.

	\item Happiness is attributed to different values of the independent variables. But it is the only emotion that is highly perceived when the linear velocity is $0$ with four over ten. However the oscillation angle for these four cases is equally divided between the values $0.174$ and $3.49$. %TODO BTW, is this angle in radians? It is a quite big difference...
So it seems that happiness is mainly attributed to fast angular velocities and big oscillations angles. Specific agreement among the other features is not present.

	\item Anger is highly perceived with an angle of $0.087$ (seven over ten) and with linear velocities over $200 mm/s$. It seems that people attribute anger to fast velocities, both angular and linear, small angle of oscillation and the robot facing the persons while approaching them. 
\end{itemize} 
%alpha general

>>>>

A table with subjects' answers was generated for each treatment. Table~\ref{table:example_table} is an example of the generated tables.

@@

For each table, mean, standard deviation, and median were calculated. It was not possible to use ANOVA test over the data because the assumption of normality is not satisfied in the collected data. This was checked using the Shapiro-Wilk Test.
Additionally, a contingency table for each emotion was generated in each treatment, as it is depicted in Table~\ref{table:table_contingency}, where the intensity for the rest of emotions is calculated as their mean of their intensity, including the option ``other''. Therefore a total of seven tables for each treatment were created. For all these tables, Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$) was calculated, which is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (e.g., $Kappa$), $\alpha$ is a generalization of several known reliability indices, and it applies to any number of observers, any number of categories, any type of data, incomplete or missing data, large and small sample sizes.

This calculation was done using the R's package \textit{irr}. To improve contingency tables' interpretation, it was decided to just register emotions' alpha and intensity values greater than zero. The $\alpha$ for each treatment is shown in Figure~\ref{fig:alpha_experience}. As it could be observed, this table does not provide any information about the interpretation for each emotion. Therefore, it was decided to create top ten raking of treatments from the contingency tables of each emotion. This ranking is generated considering the following items in the respective order:(i) the mean intensity of the respective emotion, (ii) the alpha agreement of contingency table for the respective emotion, and (iii) the alpha agreement for the specific treatment. From the resulting tables~\ref{table:happy_top_ten},~\ref{table:Angry_top_ten},~\ref{table:Sad_top_ten}, and~\ref{table:Scared_top_ten}, it is possible to notice:

@@

	\item \textit{Fear} was the only emotion that had six over ten movements obtaining both general and specific alpha agreement over 0.41, which is the lower bound for moderate agreement~\cite{Viera2005}. Also, it was mostly selected (eight over ten) when $direction = \pi$ and $orientation = \pi$. The angle and the angular velocity attributed six over ten times were $0.087$ and $2$, respectively. So it seems that people perceive the movement as a fear expression when the robot is looking at them and moving far from them fast.

	\item \textit{Sadness} was mostly attributed to linear velocities of $0 mm/s$ (two over ten) and $200 mm/s$ (eight over ten), and angular velocities of $0 rad/s$ and $1 rad/s$, with higher predilection of the second (eight over ten). It seems that people attribute sadness to slow velocities with slow angular velocity and small oscillation angle. Regarding the other two features, there is not any defined pattern that could lead to make a generalization.

	\item \textit{Happiness} is attributed to different values of the independent variables. But it is the only emotion that is highly perceived when the linear velocity is $0$ with four over ten. However the oscillation angle for these four cases is equally divided between the values $0.174$ and $3.49$. %TODO BTW, is this angle in radians? It is a quite big difference...
So it seems that happiness is mainly attributed to fast angular velocities and big oscillations angles. Specific agreement among the other features is not present.

	\item \textit{Anger} is highly perceived with an angle of $0.087$ (seven over ten) and with linear velocities over $200 mm/s$. It seems that people attribute anger to fast velocities, both angular and linear, small angle of oscillation and the robot facing the persons while approaching them. 
\end{itemize} 
%alpha general
